Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
pf <- read.csv("pseudo_facebook.tsv", sep = "\t")
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)
qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks = 1:31)
```

***

#### What are some things that you notice about this histogram?
Response: Unusually high amount of users with dob_day = 1

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response: Marriage

#### How many of your friends do you think saw that post?
Response: 150

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response: 33%

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}
qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month)
```

#### Let???s take another look at our plot. What stands out to you here?
Response: Anomaly only visible in January. Maybe users disguise real dob_day with 01.01.YYYY?

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira???s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x = friend_count, data = pf)
```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(x = friend_count, data = pf, binwidth = 25) +
  scale_x_continuous(limits = c(0,1000), breaks = seq(0, 1000, 50))
```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response: women

#### What's the difference between the median friend count for women and men?
Response: 22

#### Why would the median be a better measure than the mean?
Response: because the distribution of friend counts is skewed

***

### Tenure
Notes:

```{r Tenure}

```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(x = tenure/365, data = pf, binwidth = 0.25, color = I("black"), fill = I("black")) +
  scale_x_continuous(breaks = seq(1 ,8 ,1), limits = c(0,8))
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
qplot(x = tenure/365, data = pf, binwidth = 0.25, xlab = "Tenure in years", ylab = "Number of users in sample") +
  scale_x_continuous(breaks = seq(1 ,8 ,1), limits = c(0,8))
```

***

### User Ages
Notes:

```{r User Ages}
qplot(x = age, data = pf,
      xlab = "Age in years",
      ylab = "Number of users in sample",
      binwidth = 1) +
  scale_x_continuous(breaks = seq(13, max(pf$age), 5), limits = c(13, max(pf$age)))
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***


### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:


***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
plt_fc <- qplot(x = friend_count, data = pf)
plt_fc_log <- plt_fc + scale_x_log10()
plt_fc_sqrt <- plt_fc + scale_x_sqrt()
grid.arrange(plt_fc, plt_fc_log, plt_fc_sqrt)
```

***


### Frequency Polygons

```{r Frequency Polygons}
qplot(x = www_likes,
      data = subset(pf, !is.na(gender)),
      geom = "freqpoly",
      color = gender,
      xlab = "Number of likes in WWW",
      ylab = "Percentage of users with resp. friend count") +
  scale_x_continuous() +
  scale_x_log10()
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
by(pf$www_likes, pf$gender, sum)
```


***

### Box Plots
Notes:

```{r Box Plots}

```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
qplot(x = gender, y = friend_count, data = subset(pf, !is.na(gender)), geom = "boxplot") +
  coord_cartesian(ylim = c(0,1000))  
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
qplot(x = gender, y = friendships_initiated, data = subset(pf, !is.na(gender)), geom = "boxplot") +
  coord_cartesian(ylim = c(0, 150))

by(pf$friendships_initiated, pf$gender, summary)
```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
pf$mobile_checkin <- NA
pf$mobile_checkin <- ifelse(pf$mobile_likes > 0, TRUE, FALSE)
pf$mobile_checkin <- factor(pf$mobile_checkin)

summary(pf$mobile_checkin)
sum(pf$mobile_checkin == TRUE) / length(pf$mobile_checkin)
```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!
