Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv("../lesson3/pseudo_facebook.tsv", sep = "\t")
qplot(x = age, y = friend_count, data = pf)
```

***

#### What are some things that you notice right away?
Response: 
Younger users seem to have more friends, although there are some very old users with a lot friends too

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf) + geom_point() +
  xlim(13, 90)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) + geom_point(alpha = 1/20) +
  xlim(13, 90)
```

#### What do you notice in the plot?
Response:
Most users of all ages have friend counts well below 1000
***

### Coord_trans()
Notes:

```{r Coord_trans()}

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20) +
  xlim(13, 90) + 
  coord_trans(y = "sqrt")
```

#### What do you notice?
The plot now resembles the data better, as extreme values for friends_count look more dispersed
***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
  coord_trans(y = "sqrt") +
  xlim(13, 90)
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)

pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age)
```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) +
  geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 0.05,
             position = position_jitter(h = 0),
             color = "grey") +
  coord_trans(y = "sqrt") +
  geom_line(stat = "summary", fun.y = mean, color = "red") +
  geom_line(stat = "summary", fun.y = quantile,
            fun.args = list(probs = 0.1), linetype = 2, color = "orange") +
  geom_line(stat = "summary", fun.y = quantile,
            fun.args = list(probs = 0.9), linetype = 2, color = "orange") +
  geom_line(stat = "summary", fun.y = median, color = "blue") +
  coord_cartesian(xlim = c(13, 90), ylim = c(0, 1000))
```

#### What are some of your observations of the plot?
Response:
Users having >= 1000 friends are rare.
Data for users with age >= 75 seems to be biased
***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(x = pf$age, y = pf$friend_count, method = "pearson")
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  coord_cartesian(xlim = c(0, quantile(pf$www_likes_received, 0.95)),
                  ylim = c(0, quantile(pf$likes_received, 0.95)))
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
cor.test(pf$likes_received, pf$www_likes_received)
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
library(alr3)
data(Mitchell)
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x = Month, y = Temp), data = Mitchell) + 
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(x = (Month%%12), y = Temp), data = Mitchell) + 
  geom_point()
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

```

***

### Age with Months Means

```{r Age with Months Means}
library(tidyverse)
pf <- read.delim("../lesson3/pseudo_facebook.tsv")
pf$age_with_months <- (pf$age + (1 - pf$dob_month / 12)) 
```

Programming Assignment
```{r Programming Assignment}
pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age_with_months)
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(data = subset(pf.fc_by_age_months, age_with_months < 71), aes(x = age_with_months, y = friend_count_mean )) +
  geom_line()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

