---
output:
  html_document: default
  pdf_document: default
---
# White wine exploration by Stefan Langenbach

## Udacity data analyst nanodegree

```{r echo=FALSE, message=FALSE, warning=FALSE, setup}
# use MRO repo from end of January
options(repos = c(
  CRAN = "https://mran.revolutionanalytics.com/snapshot/2017-01-27"))

# load required packages
library(ggplot2)
library(gridExtra)
library(dplyr)
library(GGally)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, load_data}
# load data
wine <- read.csv("./wineQualityWhites.csv")
```

The dataset we are going to investigate consists of roughly 4.900 records of
white wines and 12 attributes specifying their (chemical) properties,
i.e. acidity, pH-level, alcohol-percentage and quality rating. Each wine
was rated by at least 3 experts in order to provide a quality score. The score
ranges from 0 (worst) to 10 (best).

```{r echo=FALSE, message=FALSE, warning=FALSE, structure}
str(wine)
```

The dataset does not contain any missing values.

```{r echo=FALSE, message=FALSE, warning=FALSE, na_values}
sum(is.na.data.frame(wine))
```

Investigating summary statistics for our dataset, it is obvious that some of the
variables contain extreme values. For example: The maximum value for residual
sugar is roughly 12.5 times the median. Other variables containing extreme 
values are chlorides (maximum value is roughly 8 times median value), and 
free.sulfur.dioxide. Another interesting insight is that no wine reached 
the maximum quality rating of 10.

```{r echo=FALSE, message=FALSE, warning=FALSE, summary}
summary(wine)
```

# Univariate Plots Section

Let's move on with a visual investigation of our dataset. The following plots
will investigate each variable individually.

## Non-chemical variables
```{r echo=FALSE, message=FALSE, warning=FALSE, u_quality}
ggplot(wine, aes(as.factor(quality))) +
  geom_bar() + 
  labs(title = "Wines by quality",
       x = "quality",
       y = "Number of wines")
```

Wine rating ranges from 3 to 9 and most wines were given a rating of 6. There
are only very few wines with particularly bad (quality = 3) or 
good (quality = 9) ratings.

```{r echo=FALSE, message=FALSE, warning=FALSE, u_alcohol} 
ggplot(wine, aes(x = alcohol)) +
  geom_histogram(aes(fill = ..count..), show.legend = F) +
  scale_x_continuous(breaks = c(0, 8, 9, 9.5, 10, 10.5, 11, 11.5, 12, 12.5, 13), 
                     minor_breaks = seq(9, 12.5, 0.25)) +
  labs(title = "Wines by alcohol percentage",
       x = "alcohol (% by volume)", 
       y = "Number of wines")
```
```{r echo=FALSE, message=FALSE, warning=FALSE, summary_alcohol}
summary(wine$alcohol)
```

The distribution of percentage of alcohol of wines in the dataset is multi-modal.
Although most wines have between 9 and 9.5% of alcohol, there are more peaks 10.5,
close to 11 and close to 12.5.

```{r echo=FALSE, message=FALSE, warning=FALSE, u_alc_level}
wine <- mutate(wine, alc_level = cut(wine$alcohol, breaks = 4))

ggplot(wine, aes(x = alc_level)) +
  geom_bar() +
  labs(title = "Wines by alcohol intervals", 
       x = "alcohol intervals (% of volume)",
       y = "Number of wines")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, u_sulphates} 
ggplot(wine, aes(x = sulphates)) +
  geom_freqpoly() +
  geom_vline(aes(xintercept = median(wine$sulphates), colour = "red"), 
             show.legend = F) +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  labs(title = "Wines by sulphates", 
       caption = "Vertical red line indicates median of sulphates", 
       x = expression(sulphates~(g/dm^3)), 
       y = "Number of wines")
```
```{r echo=FALSE, message=FALSE, warning=FALSE, summary_sulphates}
summary(wine$sulphates)
```

Interestingly, sulphates has a bi-modal distribution with peaks at about 0.375 
and 0.475. Furthermore there are few wines with high values of sulphates
(above 0.85).

```{r echo=FALSE, message=FALSE, warning=FALSE, u_sulphate_level}
wine <- mutate(wine, sul_level = cut(wine$sulphates, breaks = 4))

ggplot(wine, aes(sul_level)) +
  geom_bar() +
  labs(title = "Wine by sulphate intervals", 
       x = "sulphate intervals", 
       y = "Number of wines")
```


```{r echo=FALSE, message=FALSE, warning=FALSE, u_ph} 
ggplot(wine, aes(x = pH)) +
  geom_freqpoly() +
  geom_vline(aes(xintercept = median(wine$pH), colour = "red"),
             show.legend = F) +
  scale_x_continuous(breaks = seq(2, 4, 0.1)) +
  labs(title = "Wines by pH-value",
       caption = "Vertical red lines indicates median of pH",
       y = "Number of wines")
```

pH values seem to be almost normally distributed with a peak at 3.15. The
distribution has a longer tail at the right side.

```{r echo=FALSE, message=FALSE, warning=FALSE, u_residual_sugar}
ggplot(wine, aes(x = residual.sugar)) +
  geom_freqpoly(binwidth = 0.25) +
  scale_x_continuous(breaks = c(seq(0, 5, 1), seq(5, 80, 10)), 
                     minor_breaks = seq(0, 5, 1)) +
  labs(title = "Wines by residual sugar",
       x = expression(residual~sugar~(g/dm^3)),
       y = "Number of wines")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, u_residual.sugar2} 
ggplot(wine, aes(x = residual.sugar)) +
  geom_freqpoly() +
  scale_x_log10(breaks = c(1, 1.75, 8, 14, 100), 
                minor_breaks = seq(0, 15, 1)) +
  labs(title = "Wines by residual sugar", 
       x = "log10 of residual sugar", 
       y = "Number of wines")
```
```{r echo=FALSE, message=FALSE, warning=FALSE, summary_residual.sugar} 
summary(wine$residual.sugar)
```

As already observed during investigation of summary statistics, residual.sugar
contains extreme values, thus skewing standard plots without scaling of
the x-axis. The plot above uses a log10-scaled x-axis and provides better 
insightinto the distribution of residual sugar. It appears to be multi-modal 
with peaks at roughly 1.75, 8 and 14.

```{r echo=FALSE, message=FALSE, warning=FALSE, u_sugar_intervals}
wine <- mutate(wine, sugar_level = cut(wine$residual.sugar, breaks = 16))

ggplot(wine, aes(x = sugar_level)) +
  geom_bar() +
  labs(title = "Wines by residual sugar intervals in grams",
       x = "residual sugar intervals",
       y = "Number of wines")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, u_density} 
ggplot(wine, aes(x = density)) +
  geom_freqpoly() +
  geom_vline(aes(xintercept = median(wine$density), colour = "red"), 
             show.legend = F) + 
  scale_x_continuous(breaks = seq(0.9, 1.1, 0.01)) +
  labs(title = "Wines by density", 
       caption = "Vertical red line indicates median of density", 
       x = expression(density~(g/cm^3)), 
       y = "Number of wines")
```

Most wines have a density close to 1.

## Chemical variables
Let's also investigate the remaining chemical variables:

* fixed.acidity
* volatile.acidity
* citric.acid
* chlorides
* free.sulfur.dioxide
* total.sulfur.dioxide

```{r echo=FALSE, message=FALSE, warning=FALSE, u_fixed.acidity} 
ggplot(wine, aes(x = fixed.acidity)) +
  geom_histogram(aes(fill = ..count..), show.legend = F) +
  scale_x_continuous(breaks = c(0, 3, 6, 6.5, 7, 7.5, 9, 12, 15), 
                     minor_breaks = seq(6, 7.5, 0.25)) +
  labs(title = "Wines by fixed acidity", 
       x = expression(fixed~acidity~(g/dm^3)))
```

For most wines fixed acidity lies between 6 and 7.25 with two peaks at 6.5 and
7.25

```{r echo=FALSE, message=FALSE, warning=FALSE, u_volatile.acidity} 
ggplot(wine, aes(x = volatile.acidity)) +
  geom_histogram(aes(fill = ..count..), binwidth = 0.05, show.legend = F) +
  scale_x_continuous(breaks = c(0.1, 0.2, 0.3, 0.6, 0.9), 
                     minor_breaks = seq(0.15, 0.3, 0.025)) +
  labs(title = "Wines by volatile acidity", 
       x = expression(volatile~acidity~(g/dm^3)), 
       y = "Number of wines")
```

Volatile acidity peaks at roughly 0.25 for most wines, but there are two additional
plateaus at about 0.2 and 0.3.

```{r echo=FALSE, message=FALSE, warning=FALSE, u_citric.acid} 
ggplot(wine, aes(x = citric.acid)) +
  geom_histogram(aes(fill = ..count..), binwidth = 0.05, show.legend = F) +
  scale_x_continuous(breaks = c(0, 0.25, 0.35, 0,5, 1, 1.5), 
                     minor_breaks = c(0.25, 0.3, 0.35)) +
  labs(title = "Wines by citric acid", 
       x = expression(citric~acid~(g/dm^3)), 
       y = "Number of wines")
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(wine$citric.acid)
```

Majority of wines have a concentration of citric acid ranging from 0.25 to 0.35.
The maximum value is 1.66, which is roughly 5 times the median.

```{r echo=FALSE, message=FALSE, u_chlorides} 
ggplot(wine, aes(x = chlorides)) +
  geom_freqpoly() +
  scale_x_continuous(breaks = seq(0, 0.5, 0.05)) +
  labs(title = "Wines by chlorides", 
       x = expression(chlorides~(g/dm^3)), 
       y = "Number of wines")
```


```{r echo=FALSE, message=FALSE, u_chlorides2} 
ggplot(wine, aes(x = chlorides)) +
  geom_freqpoly() +
  scale_x_log10(breaks = c(0, 0.01, 0,02, 0.03, 0.04, 0.05, 0.06, 0.1, 0.25),
                minor_breaks = seq(0, 0.1, 0.01)) +
  labs(title = "Wines by chlorides", 
       x = "log10 of chlorides", 
       y = "Number of wines")
```
```{r echo=FALSE, message=FALSE, warning=FALSE, summary_chlorides}
summary(wine$chlorides)
```
Similar to residual.sugar, chlorides contains extreme values, thus 
skewing standard plots without scaling of the x-axis.
The plot above uses a log10-scaled x-axis and provides better insight
into the distribution of chlorides. Most wines have concentration of chlorides
between 0.03 and 0.06.

```{r echo=FALSE, message=FALSE, warning=FALSE, u_free.sulfur.dioxide} 
ggplot(wine, aes(x = free.sulfur.dioxide)) +
  geom_freqpoly() +
  scale_x_continuous(breaks = seq(0, 300, 50)) +
  labs(title = "Wines by free sulfur dioxide", 
       x = expression(free~sulfur~dioxide~(mg/dm^3)), 
       y = "Number of wines")
```


```{r echo=FALSE, message=FALSE, warning=FALSE, u_free.sulfur.dioxide2} 
ggplot(wine, aes(x = free.sulfur.dioxide)) +
  geom_freqpoly() + 
  scale_x_log10(breaks = c(0, 10, 25, 50, 75, 100)) +
  labs(title = "Wines by free sulfur dioxide", 
       x = "log10 of free sulfur dioxide", 
       y = "Number of wines")
```
```{r echo=FALSE, message=FALSE, warning=FALSE, summary_free_sulfur_dioxide}
summary(wine$free.sulfur.dioxide)
```

Also free sulfur dioxide contains extreme values and thus the x-axis has been
scaled to log10. Most wines have concentration of free sulfur dioxides between
25 and 50.

```{r echo=FALSE, message=FALSE, warning=FALSE, u_total.sulfur.dioxide} 
ggplot(wine, aes(x = total.sulfur.dioxide)) +
  geom_histogram(aes(fill = ..count..), binwidth = 25, show.legend = F) +
  scale_x_continuous(breaks = seq(0, 450, 50)) +
  labs(title = "Wines by total sulfur dioxide", 
       x = expression(total~sulfur~dioxide~(mg/dm^3)), 
       y = "Number of wines")
```
```{r echo=FALSE, message=FALSE, warning=FALSE, summary_total_sulfur_dioxide}
summary(wine$total.sulfur.dioxide)
```

Most wines contain between 100 and 150mg of sulfur dioxide. There is one
wine with a value three more than 3 time the median, i.e. 440.

# Univariate Analysis

### What is the structure of your dataset?
The dataset contains roughly 4.900 observations and 12 variables (X is just an
index). The variable quality is an ordered factor ranging from 0 (worst) to 
10 (best).

Additional observations:

* No wine has a quality rating below 3, none has one above 9
* Most wines contain between 9 and 11% of alcohol
* Median for residual sugar is 5.2g, but there is one wine with 65g
* Sulphates has a bi-modal distribution

### What is/are the main feature(s) of interest in your dataset?
The main features to predict quality should be alcohol, residual sugar
and sulphates.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
A combination of pure chemical features, i.e. free sulfur dioxide, chlorides and
some measure of acidity, could also have an effect on quality

### Did you create any new variables from existing variables in the dataset?
I created factors for my main variables of interest (alcohol, residual sugar and
sulphates) using cut with break points

### Of the features you investigated, were there any unusual distributions? \
Some of the variables had bi- or multi-modal distributions, i.e. sulphates and
residual sugar. In order to get an better overview about the distribution I
transformed the x-axis of their plots to log10.

# Bivariate Plots Section and analysis

Before investigating relationships among the main features of the dataset and
wine quality, let us check correlations among all variables:

```{r echo=FALSE, message=FALSE, warning=FALSE}
# omit X and interval variables created in previous analysis
tmpdata <- subset(wine, select = -c(X, alc_level, sul_level, sugar_level))

# plot correlation matrix
ggcorr(tmpdata, hjust = 1, size = 3, label = TRUE, label_size = 3,
       label_alpha = TRUE, layout.exp = 2) +
  labs(title = "Pearson correlation matrix")
```

Apparently there exist only a moderately strong positive correlation between
quality and alcohol. All other variables seem to be either unrelated or
moderately negatively correlated (density) with quality. Consequently I will
limit my bivariate analysis to relationships between

* quality and alcohol
* quality and density
* density and alcohol
* density and residual sugar

```{r echo=FALSE, message=FALSE, warning=FALSE, bar_alcohol_quality}
ggplot(wine, aes(as.factor(quality))) +
  geom_bar(aes(fill = alc_level)) +
  theme(legend.position = "bottom") +
  labs(title = "Wine quality by alcohol concentration", 
       x = "Quality", 
       y = "Number of wines", 
       fill = "Alcohol concentration intervals (% by volume)")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, box_alcohol_quality}
ggplot(wine, aes(y = alcohol, x = as.factor(quality))) +
  geom_boxplot(outlier.colour = "red", outlier.alpha = 0.1) +
  labs(title = "Alcohol vs. Quality",
       x = "quality",
       y = "alcohol (% by volume)")
```
```{r echo=FALSE, message=FALSE, warning=FALSE, scatter_alcohol_quality}
ggplot(wine, aes(x = quality, y = alcohol)) +
  geom_jitter(alpha = 0.1) +
  geom_smooth(method = "lm", se = TRUE, aes(colour = "red"), 
              show.legend = FALSE) +
  scale_x_continuous(breaks = seq(1, 9, 1)) +
  labs(title = "Alcohol vs Quality", 
       caption = "Positive relationship between variables",
       y = "alcohol (% by volume)")
```

Generally we can observe a moderate trend of rising levels of alcohol
concentration given increasing quality ratings. Although alcohol concentration
decreases when moving from quality ratings of 3 to those of 4 and 5, the
existence of outliers for these quality ratings indicates that there exists
a slight positive trend.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(wine, aes(y = density, x = as.factor(quality))) +
  geom_boxplot(outlier.colour = "red", alpha = 0.1) +
  # handle effect on outlier by zooming into plot
  scale_y_continuous(limits = quantile(wine$density, c(0.01, 0.99))) +
  labs(title = "Density vs Quality", 
       x = "quality", 
       y = expression(density~(g/cm^3)))
```

The further quality ratings increase, the lower density becomes.

```{r echo=FALSE, message=FALSE, warning=FALSE}
p1 <- ggplot(wine, aes(x = alcohol, y = density)) +
    geom_jitter(alpha = 0.1) +
    geom_smooth(method = "lm", se = FALSE) +
    labs(title = "Density vs Alcohol", 
         caption = "Negative correlation between variables",
         x = "alcohol (% by volume)",
         y = expression(density~(g/cm^3)))

p2 <- ggplot(wine, aes(x = residual.sugar, y = density)) +
    geom_jitter(alpha = 0.1) +
    geom_smooth(method = "lm", se = FALSE, aes(colour = "red"), 
                show.legend = FALSE) +
    labs(title = "Density vs. Residual sugar",
         x = expression(residual~sugar~(g/cm^3)), 
         caption = "Positive correlation between variables", 
         y = expression(density~(g/cm^3)))

# arrange plots side by side
grid.arrange(p1, p2, ncol = 2)
```

Both scatter plots visualize what has been revealed by the correlation matrix:
Strong negative/positive relationships between density and alcohol/residual
sugar

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?
As the correlation matrix revealed, only quality and alcohol have a moderate
positive relationship.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?
There is a strong positive relationship between density and residual sugar, as
well as a strong negative relationship between density and alcohol. The latter
makes sense to me, given the explanation from wineQualityInfo.txt:
"the density of wine is close to that of water depending on the percent alcohol
and sugar content". Therefore, as alcohol concentration increases, density
decreases.

### What was the strongest relationship you found?
As mentioned above, the strongest correlations (+/- 0.8) are those between
density and alcohol and density and residual.

# Multivariate Plots Section

Let's have a look at the relationship between alcohol, quality and density:

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}
ggplot(wine, aes(x = factor(quality), y = alcohol)) +
  geom_jitter(aes(colour = density), alpha = 0.1) +
  labs(title = "Alcohol vs Quality by density",
       x = "quality", 
       y = "alcohol (% by volume")
```

If we inspect the plot closely we do see higher concentrations of alcohol
associated with higher quality rating and lower values for density.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(wine, aes(x = factor(quality), y = alcohol)) +
  geom_jitter(aes(colour = residual.sugar), alpha = 0.1) +
  labs(title = "Alcohol vs Quality by residual sugar", 
       x = "quality", 
       y = "alcohol (% by volume)")
```

# Multivariate Analysis

## Linear models

Model is based on the assumptions I outlined in the univariate plots section,
i.e. that alcohol, sulphates and residual sugar would be predictors of quality,
and the insights from the correlation matrix. The results show, that a model 
solely based on alcohol is equally good (or worse taking into account adjusted 
R^2 values of roughly only 0.22 (lm1), 0.19 (lm2), 0.21 (lm3) and 0.19 (lm4).

```{r echo=FALSE, message=FALSE, warning=FALSE, linear_models}
# lm1 based on initial ideas and correlations shown by correlation matrix
lm1 <- lm(quality ~ 
            alcohol + sulphates + residual.sugar + density, data = wine)

# lm2 based only on correlation matrix
lm2 <- lm(quality ~ 
            alcohol + density, data = wine)

# lm3 based on levels of lm1
lm3 <- lm(quality ~ 
            alc_level + sul_level + sugar_level + density, data = wine)

# lm4 solely based on alcohol
lm4 <- lm(quality ~ 
            alcohol, data = wine)

# print summary statistics
summary(lm1)
summary(lm2)
summary(lm3)
summary(lm4)
```

## Residual plots

```{r echo=FALSE, message=FALSE, warning=FALSE, residual_plots}
# helper function to create residual plots 
create_res_plot <- function(y_var) {
  
  return(ggplot(aes(x = quality, y = y_var), data = wine) +
    geom_jitter(alpha = 0.05) +
    geom_hline(yintercept = 0, linetype = "dotted") +
    scale_x_continuous(breaks = seq(3, 9, 1)) +
    scale_y_continuous(breaks = seq(-3, 3, 1)) +
    labs(title = "Residual plot",
         x = "Predicted value of quality",
         y = "Residuals"))
}

# create plots for each model
p_res1 <- create_res_plot(resid(lm1)) + labs(subtitle = "Model 1")
p_res2 <- create_res_plot(resid(lm2)) + labs(subtitle = "Model 2")
p_res3 <- create_res_plot(resid(lm3)) + labs(subtitle = "Model 3")
p_res4 <- create_res_plot(resid(lm4)) + labs(subtitle = "Model 3")

# print plots in grid
grid.arrange(p_res1, p_res2, p_res3, p_res4, ncol = 2)
```

The plot above shows the predicted value of quality against the model error.
As all four models perform equaliy bad, the plots are very similiar.
It is evident that the model is especially bad for quality rating with few
observations.

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?
The visual representation of alcohol vs quality by density or residual sugar
did unfortunately not offer any new insights)

### Were there any interesting or surprising interactions between features?
No, as revealed by the correlation matrix, quality is only moderately strong
related to alcohol and density. Therefore investigating other variables did
not bring any new insights

### Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.
The four linear models I created a very simple and do not predict quality well.

# Final Plots and Summary

### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
ggplot(wine, aes(as.factor(quality))) +
  geom_bar(aes(fill = alc_level)) +
  theme(legend.position = "bottom") +
  labs(title = "Wine quality by alcohol concentration", 
       x = "Quality", 
       y = "Number of wines", 
       fill = "Alcohol concentration intervals (% by volume)")
```

### Description One
The majority of wines in the dataset is rated between 5 and 7, with 6 being the
most frequent rating. Wines included in average to above-average rating
categories (>= 6) tend to have a higher concentration of alcohol.

### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}
ggplot(wine, aes(y = alcohol, x = as.factor(quality))) +
  geom_boxplot(outlier.shape = 1, outlier.color = "red", outlier.alpha = 0.5) +
  labs(title = "Alcohol vs. Quality",
       x = "Quality", 
       y = "Alcohol (% by volume)")
```

### Description Two
Building on the insights of plot one, plot two clearly shows that median values
of alcohol concentration are rising with higher quality ratings. 
Rating categories 4 and 5 do include a fair amount of outliers.

### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}
ggplot(wine, aes(x = quality, y = alcohol)) +
  geom_jitter(alpha = 0.1) +
  geom_smooth(method = "lm", se = TRUE, aes(colour = "red"), 
              show.legend = FALSE) +
  scale_x_continuous(breaks = seq(1, 9, 1)) +
  labs(title = "Alcohol vs Quality", 
       caption = "Positive relationship between variables",
       y = "alcohol (% by volume)")
```

### Description Three
Using a linear model to visualize the relationship between quality and alcohol,
we can see that there really exists a positive trend. Apparently, as quality
ratings increase, so does alcohol concentration.

# Reflection
When reflecting on my analysis of the wine dataset, the following points come
to mind:

### Path of analysis
I started by examining every single variable in the dataset and formulated
hypotheses on which variables could be significant in predicting wine quality.
After calculating correlations, I found out that actually only alcohol (and
density) had some correlation with quality. Therefore I based my linear
models on these variables (and the factors I created from them).

### Challlenges

* As most of the (chemical) variables of the wine dataset were unknown to me,
before starting with data exploration, I had familiarize myself with their 
meaning
* In the beginning I used histograms for every visualization in univariate
analysis and played with bin width. I realized only later that freqpoly plots
have some advantage of histograms when the variable of interest is dispersed
over a wide range of values
* It took me some time to best deal with log10 transformed x-axis

### Successes

* The correlation matrix I created with ggcorr was very helpful for further
analysis.
* I found geom_smooth() to be very helpful in visualizing trends in the data.

### Surprising observations

* Before starting my analysis, that the dataset would contain more variables
that were highly correlated with wine quality. It was a bit disappointing to
have only alcohol to play with.

### Recommendations for future work

* It would be great to incorporate further (economic) variables (price, quantity
sold, origin of wine, etc.) into the dataset to investigate quality from a
different angle.

# Resources

* [R documentation](https://www.r-project.org/other-docs.html)
* [ggplot2 documentation](http://ggplot2.tidyverse.org/reference/)
* [GGally documentation](http://ggobi.github.io/ggally/#ggally)
* [tidyverse documentation](http://tidyverse.org/)